Interacting with Modal Logics \\ in the Coq Proof Assistant

### 2015-01-15 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Compress bibliography
Conflicts:
	Papers/InProgress/ModalLogicCoq/ModalCoq.pdf

Nevertheless, general automated reasoning support for modal logics is still
not as well-developed as for classical logics. Deduction tools
for modal logics are often limited to propositional, quantifier-free,
fragments or tailored to particular
modal logics and their applications \cite{CoLoSS1,CoLoSS2,ConditionalTableaux,COOL,ModalProvers,Coalescing}; 
first-order automated deduction techniques based on tableaux, sequent
calculi and connection calculi have only recently been generalized 
and implemented in a few new provers able to directly cope with
modalities \cite{JensOtten}.

Another recently explored possibility is the embedding of first-order
and even higher-order modal logics into classical higher-order logics
\cite{J23,B9,J22}, for which existing higher-order automated theorem
provers \cite{LEO-II,Satallax} exist. The embedding approach is
flexible, because various modal logics (even with multiple modalities
or varying/cumulative domain quantifiers) can be easily supported by
stating their characteristic axioms. Moreover, the approach is
relatively simple to implement, because it does not require any
modification in the source code of the higher-order prover. The prover
can be used as is, and only the input files provided to the prover
must be especially encoded. Furthermore, the efficacy and efficiency
of the embedding approach has been confirmed in philosophically
relevant benchmarks \cite{FormalTheologyRepository}. These qualities
make embedding a convenient approach for \emph{fully automated}
reasoning.

@@

Attempts to prove the existence (or non-existence) of God by means of
abstract ontological arguments are an old tradition in philosophy and
theology. G\"{o}del's proof \cite{Goedel1970,GoedelNotes} is a modern
culmination of this tradition, following particularly the footsteps of
Leibniz. Various slightly different versions of axioms and definitions
have been considered by G\"{o}del and by several philosophers who
commented on his proof (cf.
\cite{sobel2004logic,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
The formalization shown in this Section aims at being as similar as
possible to Dana Scott's version of the proof \cite{ScottNotes}. The
formulation and numbering of axioms, definitions and theorems is the
same as in Scott's notes. Even the \Coq proof scripts follow precisely
all the steps in Scott's notes. Scott's assertions are emphasized
below with comments. In contrast to the formalization in \Isabelle
\cite{AFP}, where automation via \texttt{Metis}
\cite{Hurd03first-orderproof} and \texttt{Sledgehammer}
\cite{Sledgehammer} using tools such as \texttt{Nitpick}
\cite{Nitpick}, \texttt{LEO-II} \cite{LEO-II} and \texttt{Satallax}
\cite{Satallax} has been successfully employed, the formalization in
\Coq used no automation. This was a deliberate choice, mainly because
it allowed a qualitative evaluation of the convenience of the
embedding approach for \emph{interactive} theorem proving. Moreover,
in order to formalize exactly Scott's version and not some arbitrary
version found automatically\footnote{The proofs found automatically
  by the above provers indeed differ from the one presented here:
  e.g., the strong S5 principle used below (and by Scott) is not needed; the
  ATP proofs only rely on axiom B.}, automation would have to be
heavily limited anyway. Furthermore, the deliberate preference for
simple tactics (mostly \emph{intro}, \emph{apply} and the modal
tactics described in Section \ref{sec:Tactics}) results in proof
scripts that closely correspond to common ND
proofs. This hopefully makes the formalization more accessible to
those who are not experts in \Coq's tactics but are nevertheless
interested in G\"odel's proof.

@@

The infrastructure that we have implemented for interactive and
automated reasoning in higher-order modal logics will clearly be
useful also outside philosophy; the range of potential applications is very wide.

@@

\bibliographystyle{plain}
\bibliography{Bibliography}

>>>>

Nevertheless, general automated reasoning support for modal logics is still
not as well-developed as for classical logics. Deduction tools
for modal logics are often limited to propositional, quantifier-free,
fragments or tailored to particular
modal logics and their applications 
%\cite{CoLoSS1,CoLoSS2,ConditionalTableaux,COOL,ModalProvers,Coalescing}; 
\cite{ModalProvers}; 
first-order automated deduction techniques based on tableaux, sequent
calculi and connection calculi have only recently been generalized 
and implemented in a few new provers able to directly cope with
modalities \cite{JensOtten}.

Another recently explored possibility is the embedding of first-order
and even higher-order modal logics into classical higher-order logics
\cite{J23,B9}, for which existing higher-order automated theorem
provers \cite{LEO-II,Satallax} exist. The embedding approach is
flexible, because various modal logics (even with multiple modalities
or varying/cumulative domain quantifiers) can be easily supported by
stating their characteristic axioms. Moreover, the approach is
relatively simple to implement, because it does not require any
modification in the source code of the higher-order prover. The prover
can be used as is, and only the input files provided to the prover
must be especially encoded. Furthermore, the efficacy and efficiency
of the embedding approach has been confirmed in philosophically
relevant benchmarks \cite{FormalTheologyRepository}. These qualities
make embedding a convenient approach for \emph{fully automated}
reasoning.

@@

Attempts to prove the existence (or non-existence) of God by means of
abstract ontological arguments are an old tradition in philosophy and
theology. G\"{o}del's proof 
%\cite{Goedel1970,GoedelNotes} 
\cite{Goedel1970} 
is a modern
culmination of this tradition, following particularly the footsteps of
Leibniz. Various slightly different versions of axioms and definitions
have been considered by G\"{o}del and by several philosophers who
commented on his proof (cf.
%\cite{sobel2004logic,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
\cite{sobel2004logic,Fitting,ContemporaryBibliography}).
The formalization shown in this Section aims at being as similar as
possible to Dana Scott's version of the proof \cite{ScottNotes}. The
formulation and numbering of axioms, definitions and theorems is the
same as in Scott's notes. Even the \Coq proof scripts follow precisely
all the steps in Scott's notes. Scott's assertions are emphasized
below with comments. In contrast to the formalization in \Isabelle
\cite{AFP}, where automation via \texttt{Metis}
%\cite{Hurd03first-orderproof} 
and \texttt{Sledgehammer}
%\cite{Sledgehammer} 
using tools such \texttt{LEO-II} \cite{LEO-II} and \texttt{Satallax}
\cite{Satallax} has been successfully employed, the formalization in
\Coq used no automation. This was a deliberate choice, mainly because
it allowed a qualitative evaluation of the convenience of the
embedding approach for \emph{interactive} theorem proving. Moreover,
in order to formalize exactly Scott's version and not some arbitrary
version found automatically\footnote{The proofs found automatically
  by the above provers indeed differ from the one presented here:
  e.g., the strong S5 principle used below (and by Scott) is not needed; the
  ATP proofs only rely on axiom B.}, automation would have to be
heavily limited anyway. Furthermore, the deliberate preference for
simple tactics (mostly \emph{intro}, \emph{apply} and the modal
tactics described in Section \ref{sec:Tactics}) results in proof
scripts that closely correspond to common ND
proofs. This hopefully makes the formalization more accessible to
those who are not experts in \Coq's tactics but are nevertheless
interested in G\"odel's proof.

@@

The infrastructure that we have implemented for interactive and
automated reasoning in higher-order modal logics is clearly
useful also outside philosophy; the range of potential applications is very wide.

@@

\footnotesize
\bibliographystyle{plain}
\bibliography{Bibliography}
